\documentclass[a4paper,10pt]{book}

\usepackage{tabularx}

%--custom commands---------------------------------------------
\newcommand{\lcommand}[1] {\texttt{\textbackslash #1}}
\newcommand{\lcomarg}[2] {\texttt{\textbackslash #1\{\emph{#2}\}}}
%--begin document----------------------------------------------

\begin{document}

\frontmatter

%--title page--------------------------------------------------
\thispagestyle{empty}

\begin{flushright}
\vspace*{1.5in}

{\huge Getting to Grips with \LaTeX}

\vspace{0.25in}

{\Large Andrew Roberts}


\vfill

\end{flushright}

%-----------------------------------------------------------------

\setcounter{tocdepth}{1}

\mainmatter
\chapter{Absolute beginners}

This chapter is aimed at getting familar with the bare bones on
\LaTeX{}. We will begin with creating the actual source \LaTeX{} file,
and then take you through how to feed this through the \LaTeX{} system
to produce quality output, such as postscript or PDF.

\section{The \LaTeX{} source}

The first thing you need to be aware of is that \LaTeX{} uses a markup
language in order to describe document structure and presentation. What
\LaTeX{} does is to convert your source text, combined with the markup,
into a high quality document. For the purpose of analogy, web pages work
in a similar way: the HTML is used to describe the document, but it is
your browser that presents it in its full glory --- with different
colours, fonts, sizes, etc.

\subsection{Hello World!}

Ok, so let us begin by deciding what we will actually get \LaTeX{} to
produce. As I said, we will produce the absolute bare minimum that is
needed in order to get some output, and so I think the well known 'Hello
World! approach will be suitable here.

\begin{enumerate}
  \item Open your favorite text-editor. If you use vim or emacs, they
  also have syntax highlighting that will help to write your files.
  \item Reproduce the following text in your editor. This is the
  \LaTeX{} source.

\begin{verbatim}
% hello.tex - Our first LaTeX example!

\documentclass{article}

\begin{document}

Hello World!

\end{document}
\end{verbatim}			
  \item Save your file as '\texttt{hello.tex}'. (Without the quotes!)
\end{enumerate}

\subsection{What does it all mean?}

\begin{description}

  \item{\texttt{\% hello.tex - Our first LaTeX example!}} The first line
  is a \emph{comment}. This is because it begins with the percent symbol
  (\%), which when \LaTeX{} sees it simply ignores the rest of the line.
  Comments are useful for humans to annotate parts of the source file.
  For example, you could put information about the author and the date,
  or whatever you wish. 
  
  \item{\texttt{\textbackslash documentclass\{article\}}} This line
  tells \LaTeX{} to use the article document class. A document class
  file defines the formatting, which in this case is a generic article
  format. The handy thing is that if you want change the appearance of
  your document, substitute article for another class file that exists.  
  
  \item{\texttt{\textbackslash begin\{document\}}} An educated guess
  would tell you that this command alerts \LaTeX{} that content of the
  document is about to commence.  Anything above this command are known
  generally to belong in the preamble.  
  
  \item{\texttt{Hello World!}} This was the only actual line containing
  real content --- the text that we wanted displayed on the page.
  
  \item{\texttt{\textbackslash end\{document\}}} Once again, this is not
  too difficult to understand. It tells \LaTeX{} that the document
  source is complete.  
  
\end{description}

You should also notice that each of the \LaTeX{} commands begin with a
backslash (\texttt{\textbackslash}). This is \LaTeX{}'s way of knowing
that whenever it sees a backslash, to expect some commands. Comments are
not classed as a command, since all they tell \LaTeX{} is to ignore the
line. Comments never affect the output of the document.

Note, if you want to use the backslash or percent symbols within your
text, you need to actually issue a command to tell \LaTeX{} to draw the
desired symbols, otherwise it will expect a command or a comment! The
commands are:

\begin{tabular}{ c c }
Symbol & Command \\
\% & \texttt{\textbackslash \%} \\
\textbackslash & \texttt{\textbackslash textbackslash} \\
\end{tabular}

\section{Generating the document}

It is clearly not going to be the most exciting document you have ever
seen, but we want to see it nontheless. I am assuming that you are at a
command prompt, already in the directory where \texttt{hello.tex} is
stored.

\begin{enumerate}
  \item Type the command: \texttt{latex hello} (the \texttt{.tex}
  extension is not required, although you can include it if you wish.)

  \item Various bits of info about latex and its progress will be
  displayed. If all went well, the last two lines displayed in the
  console will be:

  \begin{verbatim}
Output written on hello.dvi (1 page, 232 bytes).
Transcript written on hello.log.
\end{verbatim}

\end{enumerate}

This means that your source file has been processed and the resulting
document is called hello.dvi, which takes up 1 page and 232 bytes of
space.

Note, in this instance, due to the simplicity of the file, you only need
to run the \LaTeX{} command once. However, if you begin to create
complex documents, including bibliographies and cross-references, etc.,
\LaTeX{} needs to be executed multiple times to resolve the references.
But this will be discussed in the future when it comes up.  

\section{Viewing the document}

\LaTeX{} has now done its job, so we can view the output. The default
format is DVI (device independent), of which viewers exist freely for
most platforms.  However, the chances are that you would prefer to have
a postscript file or PDF. Fortunately, there exist tools that can
convert DVI to PS (and PDF) easily.

\subsection{Converting to Postscript}

Type the command: \texttt{dvips hello.dvi -o hello.ps}

\textsf{dvips} is the utility that actually performs the conversion. The
first argument is the DVI file to be converted. The \texttt{-o} argument
says that you want the output to be saved as a file. And the argument
immediately after is the name you wish to call it. You could give it any
name, but it makes sense to stick with hello, as well as giving it an
informative \texttt{.ps} extension.

\subsection{Converting to PDF}

There are two easy routes to get a PDF:

\begin{enumerate}
  \item Type the command: \texttt{dvipdf hello.dvi hello.pdf} (Note that there
  is no \texttt{-o} with this command, because although the utilities look
  almost identical, they have slighly differing syntax)
  \item If you already have a postscript version, then type: \texttt{ps2pdf
  hello.ps hello.pdf}
\end{enumerate}

Now it is simply a matter of using your preferred PS or PDF viewer to
see the output. What you should see at the top left of the page are the
words Hello World! and at the bottom is the current page number. All in
a standard times font.

\section{Summary}

Ok, we've created possibly the simpliest possible document that \LaTeX{}
will produce (except for a blank page of course!) which is why it is not
much to look at. However, now we have seen the basics, and how to
actually use the \LaTeX{} software, we can progress towards the more
typical documents that you are likely to produce.

\chapter{Document structure}

This chapter progresses significantly from the previous --- very
simplistic --- chapter. The goal is to produce a fairly basic article, of
similar style to what a research paper would resemble. To achieve this
efficiently, we will focuse largely on document structure.

\LaTeX{} practically forces you to declare structure within your documents.
This is a good thing though. Because once \LaTeX{} understands how you want
your document organised, it will take care of all the tedious business
of the layout and presentation for you. The separation of content and
layout allows you to concentrate on the job at hand, i.e., communicating
your ideas.

\section{Preamble}

If you recall from the previous tutorial, the preamble the everything
from the start of the \LaTeX{} source file until the
\texttt{\textbackslash begin\{document\}} command. It normally contains
commands that affect the entire document.

\begin{verbatim}
% simple.tex - A simple article to illustrate document structure.

\documentclass{article} 
\usepackage{times}

\begin{document}
\end{verbatim}		

The first line is a comment (as denoted by the \texttt{\%} sign). The
\texttt{\textbackslash documentclass} command takes an argument, which
in this case is \texttt{article}, because that's the type of document we
want to produce. Other default classes that exist are \texttt{book},
\texttt{report}, \texttt{letter}, etc. It is also possible to create
your own, as is often done by journal publishers, who simply provide you
with their own \emph{class} file, which tells \LaTeX{} how to format your content.
But we'll be happy with the standard article class for now! 

\texttt{\textbackslash usepackage} is an important command that tells
Latex to utilise some external macros. In this instance, I specified
\texttt{times} which means \LaTeX{} will use the Postscript Times type 1
fonts, which look nicer :) And finally, the \texttt{\textbackslash
begin\{document\}}. This
strictly isn't part of the preamble, but I'll put it here anyway, as it
implies the end of the preamble by nature of stating that the document
is now starting.

\section{Top Matter}

At the beginning of most documents will be information about the
document itself, such as the title and date, and also information about
the authors, such as name, address, email etc. All of this type of
information within Latex is collectively referred to as top matter.
Although never explicitly specified, that is, there is no such
\texttt{\textbackslash topmatter} command, you are likely to encouter
the term within Latex documentation.

An example:

\begin{verbatim}
\title{How to Structure a \LaTeX{} Document} 
\author{Andrew Roberts\\
  School of Computing,\\ 
  University of Leeds,\\ 
  Leeds,\\ 
  United Kingdom,\\
  LS2 9JT\\ 
  \texttt{andyr@comp.leeds.ac.uk}} 
\date{\today} 
\maketitle
\end{verbatim}

The \texttt{\textbackslash title} command is fairly obivous. Simply put
the title you want between the curly braces. \texttt{\textbackslash
author} would also seem easy, until you notice that I've crammed in all
sorts of other information along with the name. This is merely a common,
albeit, ungraceful hack, due to the default article class being a tad
basic. If you are provided with a class file from a publisher, or if you
use the AMS article class (amsart), then you have a more logical
approach to entering author information. In the meantime, you can see
how the new line command (\textbackslash\textbackslash) has been used so
that I could produce my address. My email address is at the end, and the
\texttt{\textbackslash texttt} command formats the email address using a monospaced font. The
\texttt{\textbackslash date} command takes an argument to signify the date the document was
written.  I've used a built-in command called
\texttt{\textbackslash today} which, when
processed by \LaTeX{}, will be replaced with the current date.  But you are
free to put whatever you want as a date, in no set order. If braces are
left empty, then the date is then omitted.  Without
\texttt{\textbackslash maketitle}, the top
matter would not appear in the document. So it is needed to commit your
article attributes to paper.  

\section{Abstract}

As most research papers have an abstract, then there is a predefined
command for telling \LaTeX{} which part of the content makes up the
abstract. This should appear in its logical order, therefore, after the
top matter, but before the main sections of the body.

\begin{verbatim}
\begin{abstract} 
Your abstract goes here...  
...  
\end{abstract}
\end{verbatim}		

\section{Sectioning commands}

The commands for inserting sections are fairly intuitive. Of course,
certain commands are appropriate to different document classes. For
example, a book has chapters but a article doesn't. Here is an edited
version of some of the structure commonds in use.

\begin{verbatim}
\section{Introduction} 
This section's content...

\section{Structure} 
This section's content...

\subsection{Top Matter} 
This subsection's content...

\subsubsection{Article Information}
This subsubsection's content...
\end{verbatim}

As you can see, the commands are fairly intuitive. Notice that you do
not need to specify section numbers. Latex will sort that out for you!
Also, for sections, you do not need to markup which content belongs to a
given block, using \lcommand{begin} and \lcommand{end} commands, for example.

\begin{table}
  \begin{tabular}{|l|l|}
    \hline
    Command & Level \\
    \lcomarg{part}{part} & -1 \\
    \lcomarg{chapter}{chapter} & 0 \\
    \lcomarg{section}{section} & 1 \\
    \lcomarg{subsection}{subsection} & 2 \\
    \lcomarg{subsubsection}{subsubsection} & 3 \\
    \lcomarg{paragraph}{paragraph} & 4 \\
    \lcomarg{subparagraph}{subparagraph} & 5 \\
    \hline
  \end{tabular}
  \caption{Possible section commands.\ref{table:sections}}
\end{table}

Numbering of the sections is performed automatically by Latex, so don't
bother adding them explicitly, just insert the heading you want between
the curly braces. If you don't want sections number, then add an
asterisk (\texttt{*}) after the section command, but before the first curly
brace, e.g., \verb|\section*{A Title Without Numbers}|.  

\section{The bibliography}

Any good research paper will have a whole list of references. In this
example document, I have included one. If you look at the PDF version,
then after the first instance of `Latex' in the introduction, you should
notice a numbered reference. And at the end of the document, you can see
the full reference.

Fortunately, Latex has a slightly more intelligent approach to managing
your references than the average word processor, like MS Word for
example, where everything has to be inputted manually (unless you
purchase a 3rd party add-on). There are two ways to insert your
references into Latex: the first is to store them in an external file
and then link them via a command to your current document, or secondly,
embed them within the document itself. In this chapter, I shall quickly
cover the latter. Although, the former will be covered in depth in a
future tutorial, as it is by far the most efficient and flexible.

There are two stages to setting up your biblography/references in a
document. The first is to set up a bibliography environment, which is
where you provide Latex with the details of the references. The second
is the actual citation of your references within your document.

The following code was used in creating the bibliography environment for
the example document in this tutorial. It is located immediately after the last
line of the document content, but before the \verb|\end{document}| command.

\begin{verbatim}
\begin{thebibliography}{9}
	\bibitem{lamport94} Leslie Lamport, \emph{\LaTeX: A Document
	Preparation System}.  Addison Wesley, Massachusetts, 2nd Edition,
	1994.
\end{thebibliography}
\end{verbatim}

Ok, so what is going on here? The first thing to notice is the establishment of
the environment. \texttt{thebibliography} is a keyword that Latex recognises as
everything between the \texttt{begin} and \texttt{end} tags as being data for
the bibliography. The optional argument which I supplied after the
\texttt{begin} statement is telling Latex how wide the item label will be when
printed\footnote{An item label is simply the number that appears before the
actual reference that allows you to cross reference it with the cited number
within the document}. Note however, that it is not a literal parameter, i.e the
number 9 in this case, but a text width. Therefore, I am effectively telling
Latex that I will only need reference labels of one character in width, which
means no more than nine references in total. If you want more than ten, then
input a two-digit number, such as '99' which permits less than 100 references.

Next is the actual reference entry itself. This is prefixed with the
\lcomarg{bibitem}{cite\_key} command. The \emph{cite\_key} is should be a unique
identifier for that particular reference, and is often some sort of
mnemonic consisting of any sequence of letters, numbers and punctuation
symbols (although not a comma). I often use the surname of the first
author, followed by the last two digits of the year (hence `\texttt{lamport94}').
If that author has produced more than one reference for a given year,
then I add letters after, `a', `b', etc. But, you should whatever works
for you. Everything after the key is the reference itself. You need to
type it as you want it to be presented. I have put the different parts
of the reference, such as author, title, etc., on different lines for
readability. These line breaks are ignored by Latex. I wanted the title to be
in italics, so I used the \lcommand{emph} command to achieve this.

To actually cite a given reference within your document is very easy. Goto the
point where you want the citation to appear, and use the following:
\verb|\cite{cite_key}|, where the \emph{cite\_key} is that of the bibitem you wish to cite.
When Latex processes the document, the citation will be cross-referenced with
the bibitems and replaced with the appropriate number citation. The advantage
here, once again, is that Latex looks after the numbering for you. If it was
totally manual, then adding or removing a reference can be a real chore, as you
would have to re-number all the citations by hand.

Of course, it may be your preference to use a different referencing
system, such as Harvard, instead of the default numerical. This will be
covered in the future, in the mean time, why not try to experiment with
the \verb|\package{Natbib}| package.  

\appendix
\chapter{History of \LaTeX}

Some interesting stuff about \LaTeX\ldots

\end{document}